---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# psymetrics

<!-- badges: start -->
<!-- badges: end -->

The goal of psymetrics is to provide tools for extracting and visualizing psychometric model fit indices. It is compatible with models created using packages like lavaan, psych, and mirt.

## Installation

You can install the development version of psymetrics from [GitHub](https://github.com/brianmsm/psymetrics) with:

``` r
# install.packages("pak")
pak::pak("brianmsm/psymetrics")
```

## Getting Fit Indices

Here is an example of how to use the psymetrics package with a model created using lavaan.

```{r}
library(psymetrics)
library(lavaan)

# Define a simple CFA model
model <- 'visual  =~ x1 + x2 + x3
          textual =~ x4 + x5 + x6
          speed   =~ x7 + x8 + x9'

# Fit the model using lavaan
fit <- cfa(model, data = HolzingerSwineford1939, estimator = "MLR")

# Extract and print fit indices
model_fit(fit)

# You can also request specific types of indices, such as 'robust'
model_fit(fit, type = "robust")

# Or specify which indices to extract
model_fit(fit, metrics = c("cfi", "tli"))
```

This example demonstrates how to extract and print various fit indices from a confirmatory factor analysis (CFA) model using psymetrics. You can choose between standard, scaled, or robust fit indices, and even specify custom sets of indices to extract.

## Comparing Fit Indices

```{r}
fit_1 <- cfa(model, data = HolzingerSwineford1939, estimator = "MLR")
fit_2 <- cfa(model, data = HolzingerSwineford1939, estimator = "ULSM")

compare_model_fit(fit_1, fit_2)
```

In this example, compare_model_fit is used to compare the fit indices of two different models. This function allows you to easily see the differences in model fit across different estimation methods or model specifications.

