---
title: "Get started with fit indices"
description: >
  A compact end-to-end workflow to fit a lavaan CFA model, extract fit indices,
  and compare two fitted models with psymetrics.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with fit indices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)

has_lavaan <- requireNamespace("lavaan", quietly = TRUE)
```

## Goal

By the end of this guide, you will be able to:

1. Fit a CFA model with `lavaan`.
2. Extract fit indices with `model_fit()`.
3. Compare alternative estimators with `compare_model_fit()`.
4. Fit an ordinal-item version with `WLSMV` and `ordered = TRUE`.

## Prerequisites

- R >= 4.1.0
- Packages: `psymetrics`, `lavaan`

```{r, eval=FALSE}
install.packages("lavaan")
```

## Setup

```{r, eval=has_lavaan}
library(psymetrics)
library(lavaan)
```

## Step 1: Fit a baseline CFA model

```{r, eval=has_lavaan}
model <- '
  visual  =~ x1 + x2 + x3
  textual =~ x4 + x5 + x6
  speed   =~ x7 + x8 + x9
'

fit_mlr <- cfa(
  model,
  data = HolzingerSwineford1939,
  estimator = "MLR"
)
```

## Step 2: Extract fit indices

```{r, eval=has_lavaan}
model_fit(fit_mlr)
```

Request only selected metrics when needed:

```{r, eval=has_lavaan}
model_fit(fit_mlr, metrics = c("cfi", "tli", "rmsea"))
```

## Step 3: Compare two fitted models

```{r, eval=has_lavaan}
fit_ulsm <- cfa(
  model,
  data = HolzingerSwineford1939,
  estimator = "ULSM"
)
compare_model_fit(MLR = fit_mlr, ULSM = fit_ulsm)
```

## Step 4: Ordinal example with `WLSMV` (`ordered = TRUE`)

This example converts the nine Holzinger-Swineford items (`x1` to `x9`) into
four ordered categories, then fits the same CFA model with `WLSMV`.

```{r, eval=has_lavaan}
item_names <- paste0("x", 1:9)
ordered_data <- HolzingerSwineford1939

ordered_data[item_names] <- lapply(
  ordered_data[item_names],
  function(x) ordered(cut(x, breaks = 4, include.lowest = TRUE))
)

fit_wlsmv <- cfa(
  model,
  data = ordered_data,
  estimator = "WLSMV",
  ordered = TRUE
)

model_fit(fit_wlsmv)
compare_model_fit(MLR = fit_mlr, WLSMV = fit_wlsmv)
```

## Common pitfalls

- When using robust estimators, metric names may map to scaled/robust variants.
- Keep model names explicit (`MLR`, `ULSM`, etc.) for clearer comparison tables.
- This ordinal recoding is only a didactic example; use substantively justified
  category definitions in real analyses.

## Next steps

- Continue with [Reporting and visualization](reporting-and-visualization.html) to
  turn fit outputs into tables and plots.
- Continue with [SEM and parameter estimates](sem-and-estimates-lavaan.html).
- See full API details in the [Reference](../reference/index.html).
