---
title: "Reporting and visualization"
description: >
  Turn psymetrics outputs into markdown/Word tables and visualize factor loadings
  for communication-ready psychometric reporting.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reporting and visualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)

has_lavaan <- requireNamespace("lavaan", quietly = TRUE)
has_ggplot2 <- requireNamespace("ggplot2", quietly = TRUE)
```

## Goal

This guide focuses on output workflows:

1. Prepare comparison tables for markdown reports.
2. Export fit tables to Word.
3. Visualize standardized factor loadings.

## Setup

```{r, eval=has_lavaan}
library(psymetrics)
library(lavaan)

model <- '
  visual  =~ x1 + x2 + x3
  textual =~ x4 + x5 + x6
  speed   =~ x7 + x8 + x9
'

fit_mlr <- cfa(
  model,
  data = HolzingerSwineford1939,
  estimator = "MLR"
)
fit_ulsm <- cfa(
  model,
  data = HolzingerSwineford1939,
  estimator = "ULSM"
)
fit_table <- compare_model_fit(MLR = fit_mlr, ULSM = fit_ulsm)
```

## Format results for markdown, HTML, and auto output

To request markdown output explicitly:

```{r, eval=FALSE}
format_results(fit_table, output = "markdown")
```

```{r, echo=FALSE, results='asis'}
md_table <- format_results(fit_table, output = "markdown")
md_table_txt <- paste(md_table, collapse = "\n")
knitr::asis_output(paste0("```markdown\n", md_table_txt, "\n```"))
```

In an HTML report/article, running that call renders a formatted table:

```{r, eval=has_lavaan, echo=FALSE}
format_results(fit_table, output = "markdown")
```

By default, `format_results(fit_table)` uses `output = "auto"` and chooses markdown or HTML according to the rendering context:

In HTML-capable contexts, you can explicitly request HTML:

```{r}
format_results(fit_table, output = "html")
```

## Export to Word

```{r, eval=FALSE}
save_table(
  fit_table,
  path = "model_fit.docx",
  orientation = "landscape"
)
```

## Visualize factor loadings

```{r, eval=has_lavaan && has_ggplot2, fig.width=6, fig.height=4.5, fig.dpi=300, fig.alt="Dot plot of standardized factor loadings by item with confidence intervals."}
plot_factor_loadings(fit_mlr)
```

## Practical notes

- `format_results()` is useful for reproducible reports.
- `save_table()` is ideal for manuscript-ready `.docx` tables.
- Use plotting as a communication layer after checking fit and parameters.

## Next steps

- Return to [Get started with fit indices](get-started-fit-indices.html).
- Explore the [Reference](../reference/index.html) for argument-level control.
