---
title: "Reporting and visualization"
description: >
  Turn psymetrics fit objects into markdown/Word tables, visualize fit indices
  with plot_model_fit(), and add complementary loading plots for
  communication-ready psychometric reporting.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reporting and visualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)

has_lavaan <- requireNamespace("lavaan", quietly = TRUE)
has_ggplot2 <- requireNamespace("ggplot2", quietly = TRUE)
```

## Goal

This guide focuses on communication-ready output workflows:

1. Format fit results for markdown and HTML reports.
2. Export fit tables to Word.
3. Visualize single-model and multi-model fit summaries with `plot_model_fit()`.
4. Use `plot_factor_loadings()` as a complementary structural view.

## Setup

```{r, eval=has_lavaan}
library(psymetrics)
library(lavaan)

model <- '
  visual  =~ x1 + x2 + x3
  textual =~ x4 + x5 + x6
  speed   =~ x7 + x8 + x9
'

fit_mlr <- cfa(
  model,
  data = HolzingerSwineford1939,
  estimator = "MLR"
)
fit_ulsm <- cfa(
  model,
  data = HolzingerSwineford1939,
  estimator = "ULSM"
)
fit_mlr_multi <- cfa(
  model,
  data = HolzingerSwineford1939,
  estimator = "MLR",
  test = c("satorra.bentler", "mean.var.adjusted")
)

single_fit <- model_fit(fit_mlr)
compared_fits <- compare_model_fit(MLR = fit_mlr, ULSM = fit_ulsm)
single_fit_multi <- model_fit(fit_mlr_multi, standard_test = TRUE)
```

## Format results for markdown, HTML, and auto output

To request markdown output explicitly:

```{r, eval=FALSE}
format_results(compared_fits, output = "markdown")
```

```{r, eval=has_lavaan, echo=FALSE, results='asis'}
md_table <- format_results(compared_fits, output = "markdown")
md_table_txt <- paste(md_table, collapse = "\n")
knitr::asis_output(paste0("```markdown\n", md_table_txt, "\n```"))
```

In an HTML report or pkgdown article, the same object can render directly as a
formatted table:

```{r, eval=has_lavaan, echo=FALSE}
format_results(compared_fits, output = "markdown")
```

By default, `format_results(compared_fits)` uses `output = "auto"` and chooses
markdown or HTML according to the rendering context.

In HTML-capable contexts, you can also request HTML explicitly:

```{r, eval=has_lavaan}
format_results(compared_fits, output = "html")
```

## Export to Word

```{r, eval=FALSE}
save_table(
  compared_fits,
  path = "model_fit.docx",
  orientation = "landscape"
)
```

## Visualize a single fitted model

`plot_model_fit()` is the public plotting entrypoint for `model_fit()` and
`compare_model_fit()` objects. For a one-row `model_fit` summary, the default
style resolves to the single-fit bullet chart.

```{r, eval=has_lavaan && has_ggplot2, fig.width=6, fig.height=5.4, fig.dpi=300, fig.alt="Bullet chart of CFI, TLI, RMSEA, and SRMR for a single fitted model."}
plot_model_fit(single_fit)
```

## Visualize a model comparison

For `compare_model_fit()` objects, the default plot is the threshold-aware dot
plot, which works well as a quick side-by-side comparison.

```{r, eval=has_lavaan && has_ggplot2, fig.width=7, fig.height=5.8, fig.dpi=300, fig.alt="Threshold-aware dot plot comparing CFI, TLI, RMSEA, and SRMR across two fitted models."}
plot_model_fit(compared_fits)
```

## Try an alternative plot style

In `v0.5.0`, the supported plot styles are `default`, `bullet`, `dots`,
`bars`, and `heatmap`. The article keeps to the main workflow, while the
[plot_model_fit reference](../reference/plot_model_fit.html) documents the full
argument surface and supported constraints.

```{r, eval=has_lavaan && has_ggplot2, fig.width=7, fig.height=6.4, fig.dpi=300, fig.alt="Grouped threshold bar chart comparing fit indices for two fitted models."}
plot_model_fit(compared_fits, type = "bars")
```

## Focus on selected metrics

If you only want to communicate part of the fit summary, pass a subset with
`metrics`.

```{r, eval=has_lavaan && has_ggplot2, fig.width=7, fig.height=5.8, fig.dpi=300, fig.alt="Grouped threshold bar chart using only CFI, TLI, and RMSEA."}
plot_model_fit(
  compared_fits,
  type = "bars",
  metrics = c("CFI", "TLI", "RMSEA")
)
```

## Work with multiple test rows

When the fit object keeps both standard and non-standard test summaries,
`plot_model_fit()` can filter which rows to show with `test_mode`. Here,
`standard_test = TRUE` keeps multiple rows in the `model_fit()` result, and
`test_mode = "primary"` reduces the view to the primary non-standard summary.

```{r, eval=has_lavaan && has_ggplot2, fig.width=6, fig.height=5.4, fig.dpi=300, fig.alt="Bullet chart for the primary non-standard fit summary selected from a multi-row model_fit object."}
plot_model_fit(single_fit_multi, test_mode = "primary")
```

## Visualize factor loadings as a complementary check

`plot_model_fit()` answers how well the model fits overall. `plot_factor_loadings()`
adds a complementary view of the measurement structure after the fit summary
looks acceptable.

```{r, eval=has_lavaan && has_ggplot2, fig.width=6, fig.height=4.5, fig.dpi=300, fig.alt="Dot plot of standardized factor loadings by item with confidence intervals."}
plot_factor_loadings(fit_mlr)
```

## Practical notes

- Use `format_results()` when you want the same fit object rendered in reports.
- Use `save_table()` when the output needs to go into a `.docx` workflow.
- Use `plot_model_fit()` for the supported `CFI`/`TLI`/`RMSEA`/`SRMR` workflows in
  `v0.5.0`, and keep the reference page nearby for argument-level detail.
- Use `plot_factor_loadings()` after fit checking when you want to discuss item
  structure rather than overall fit.

## Next steps

- Return to [Get started with fit indices](get-started-fit-indices.html).
- Continue with [SEM and parameter estimates](sem-and-estimates-lavaan.html).
- Explore the [Reference](../reference/index.html) for argument-level control.
