---
title: "SEM and parameter estimates with lavaan"
description: >
  Fit a SEM model with lavaan, inspect model fit, and extract parameter blocks
  with model_estimates() in psymetrics.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SEM and parameter estimates with lavaan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)

has_lavaan <- requireNamespace("lavaan", quietly = TRUE)
```

## Goal

This article shows how to:

1. Fit a SEM model with `lavaan::sem()`.
2. Compare fit across estimators.
3. Extract parameter estimates by component.

## Prerequisites

- Packages: `psymetrics`, `lavaan`

## Setup

```{r, eval=has_lavaan}
library(psymetrics)
library(lavaan)
```

## Fit SEM models

```{r, eval=has_lavaan}
sem_model <- '
  ind60 =~ x1 + x2 + x3
  dem60 =~ y1 + y2 + y3 + y4
  dem65 =~ y5 + y6 + y7 + y8

  dem60 ~ ind60
  dem65 ~ ind60 + dem60
'

fit_sem_mlr <- sem(
  sem_model,
  data = PoliticalDemocracy,
  estimator = "MLR"
)
fit_sem_ml <- sem(
  sem_model,
  data = PoliticalDemocracy,
  estimator = "ML"
)
```

## Inspect and compare fit

```{r, eval=has_lavaan}
model_fit(fit_sem_mlr)
compare_model_fit(MLR = fit_sem_mlr, ML = fit_sem_ml)
```

## Extract parameter estimates

All parameter blocks:

```{r, eval=FALSE}
model_estimates(fit_sem_mlr)
```

Only selected components:

```{r, eval=FALSE}
model_estimates(
  fit_sem_mlr,
  component = c("loading", "regression"),
  standardized = TRUE
)
```

## Practical notes

- Use `component` to narrow output to the section relevant for your report.
- Use `standardized = TRUE` when comparing relative magnitudes.

## Next steps

- Continue with [Reporting and visualization](reporting-and-visualization.html).
- See function-level details in the [Reference](../reference/index.html).
