---
title: "SEM, parameter estimates, and comparisons with lavaan"
description: >
  Fit CFA and SEM models with lavaan, inspect model fit, extract parameter blocks,
  and compare parameter estimates with model_estimates() and compare_model_estimates()
  in psymetrics.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SEM, parameter estimates, and comparisons with lavaan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)

has_lavaan <- requireNamespace("lavaan", quietly = TRUE)
```

## Goal

This article shows how to:

1. Fit related CFA and SEM models with `lavaan`.
2. Inspect and compare model fit.
3. Extract parameter estimates by component with `model_estimates()`.
4. Compare parameter estimates across models with `compare_model_estimates()`.
5. Customize comparison layouts with public `select` presets and advanced `table_args` overrides.

## Prerequisites

- Packages: `psymetrics`, `lavaan`

## Setup

```{r, eval=has_lavaan}
library(psymetrics)
library(lavaan)
```

## Fit related CFA and SEM models

We will use the same measurement structure in two fitted models:

- a CFA model with three latent factors;
- a SEM model that keeps the same measurement model and adds regressions among factors.

This gives us a realistic comparison where some parameters are shared across models and some appear only in the SEM.

```{r, eval=has_lavaan}
hs_cfa_model <- '
  visual =~ x1 + x2 + x3
  textual =~ x4 + x5 + x6
  speed =~ x7 + x8 + x9
'

hs_sem_model <- '
  visual =~ x1 + x2 + x3
  textual =~ x4 + x5 + x6
  speed =~ x7 + x8 + x9
  textual ~ visual
  speed ~ visual + textual
'

fit_cfa <- cfa(hs_cfa_model, data = HolzingerSwineford1939)
fit_sem <- sem(hs_sem_model, data = HolzingerSwineford1939)
```

## Inspect and compare fit

You can still use the fit-oriented helpers before moving to parameter-level output.

```{r, eval=has_lavaan}
model_fit(fit_cfa)
compare_model_fit(CFA = fit_cfa, SEM = fit_sem)
```

## Extract parameter estimates with `model_estimates()`

For a single fitted model, `model_estimates()` gives you parameter blocks such as loadings, regressions, variances, and means.

```{r, eval=has_lavaan}
model_estimates(fit_sem)
```

You can narrow the output to the parts most relevant to your workflow. For example, when you only want loadings and regressions:

```{r, eval=has_lavaan}
model_estimates(
  fit_sem,
  component = c("loading", "regression"),
  standardized = TRUE
)
```

## Compare estimates across models

The simplest comparison uses the default `select = "ci"`, which shows estimates with confidence intervals.

```{r, eval=has_lavaan}
compared_default <- compare_model_estimates(CFA = fit_cfa, SEM = fit_sem)
compared_default
```

This is especially useful when one model contains parameters that do not exist in the other. In this example, the regression rows only appear for the SEM, while shared loadings still line up across both models.

## Focus the comparison on selected components

As with `model_estimates()`, you can restrict the comparison to a subset of components.

```{r, eval=has_lavaan}
compare_model_estimates(
  CFA = fit_cfa,
  SEM = fit_sem,
  component = c("loading", "regression")
)
```

Use this when you want a tighter table for a manuscript, report, or diagnostics pass.

## Use public `select` presets

The public `select` argument controls which statistics are shown and how they are arranged by default for the object.

For example, `select = "se_p"` shows estimates, standard errors, and significance stars in a compact single column per model.

```{r, eval=has_lavaan}
compared_se <- compare_model_estimates(
  CFA = fit_cfa,
  SEM = fit_sem,
  select = "se_p"
)
compared_se
```

If you prefer numeric p-values in a second column, use a two-column preset such as `"ci_p2"`.

```{r, eval=has_lavaan}
compare_model_estimates(
  CFA = fit_cfa,
  SEM = fit_sem,
  select = "ci_p2"
)
```

## Render the same comparison in another layout

In most cases, prefer setting `select` directly in `compare_model_estimates()`. That keeps the default presentation close to where the comparison is created.

`format_results(..., table_args = list(select = ...))` is useful when you already have a comparison object and want to render the same aligned estimates in a different way for a specific output. This works because the `compare_model_estimates()` result keeps the underlying estimate, standard error, confidence interval, and p-value columns needed to rebuild the display later.

```{r, eval=has_lavaan}
format_results(
  compared_se,
  output = "text",
  table_args = list(select = "{estimate} ({ci})|{p}")
)
```

You can also use custom templates that keep everything in one cell.

```{r, eval=has_lavaan}
format_results(
  compared_se,
  output = "text",
  table_args = list(select = "{estimate}{stars} ({se})")
)
```

## When to use each layer

Start with `compare_model_estimates(select = ...)` when you know the layout you want most of the time.

Use `format_results(..., table_args = list(select = ...))` when you want an alternate console view, export, or manuscript-ready layout from the same stored comparison object without recomputing the comparison.

If you do not need that flexibility, staying with `select` inside `compare_model_estimates()` is the simpler option.

## Next steps

- Continue with [Reporting and visualization](reporting-and-visualization.html).
- See function-level details in the [Reference](../reference/index.html).

